---
title: "Improving the performance of an investment fund"
author: "Luis Moreno"
date: "12/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Report Content
1. Show performance comparison between Bitcoin, S&P 500 and Gold price.

2. Analyze Bitcoin's returns and volatility profile.
  - Could Bitcoin help improve the performance of a portfolio?
  - Could Bitcoin be used as a hedge versus inflation?

3. The CFO is looking to **lower volatility in the fund**. Explore building a portfolio using some or all of these assets. Make a recommendation that minimizes overall risk.

# Packages
```{r Required Packages, message=FALSE}
library(tidyverse)
#install.packages("quantmod")
library(quantmod)
```

# Datasets
```{r Bitcoin dataset, message=FALSE}
# Read Bitcoin data from CSV file
btc <- readr::read_csv("./bitcoin_usd.csv")

# Create a xts object for Bitcoin data
btc_xts <- as.xts(btc %>% 
                  select(open, high, low, close, volume), 
                  order.by = as.Date(btc$date))
```

```{r S&P 500 dataset, message=FALSE}
# Read S&P 500 data from CSV file
sp500 <- readr::read_csv("./sp500.csv")

# Create a xts object for S&P 500 data
sp500_xts <- as.xts(sp500 %>%
                    select(open, high, low, close, volume),
                    order.by = as.Date(sp500$date))
```

```{r Gold dataset, message=FALSE}

```

```{r CPI dataset, message=FALSE}

```

# Visualizing Price Charts
## Bitcoin Candlestick Chart
```{r Bitcoin Chart, echo=FALSE}
btc_chart <- chartSeries(to.monthly(btc_xts),
              type = "candlesticks",
              up.col = "white",
              dn.col = "black",
              log.scale = TRUE,
              theme = chartTheme("white"),
              name = "BTCUSD")
```

## S&P 500 Candlestick Chart
```{r SP500 Chart, echo=FALSE}
sp500_chart <- chartSeries(to.monthly(sp500_xts),
              type = "candlesticks",
              up.col = "white",
              dn.col = "black",
              log.scale = TRUE,
              theme = chartTheme("white"),
              name = "S&P 500")
```

```{r Visualizing Time Series, include =FALSE}
gold_lplot = 
  ggplot(monthly_data, aes(x = date, y = gold_usd)) + 
  geom_line() +
  scale_y_log10()

gold_lplot

cpi_lplot = 
  ggplot(monthly_data, aes(x = date, y = cpi_us)) +
  geom_line() +
  scale_y_log10()

cpi_lplot
```

# Visualizing Performance