---
title: "Intermediate Regression in R"
author: "Luis Moreno"
date: "11/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Required Packages}
library(fst)
library(dplyr)
library(tidyr)
library(ggplot2)
library(moderndive)
library(broom)
```

```{r Required Datasets}
taiwan_real_estate <- read_fst("/Users/luismoreno/Documents/GitHub/datacamp/datasets/taiwan_real_estate.fst")
```

# Parallel Slopes  
## Parallel Slopes Linear Regression  
- When there is **one numeric explanatory variable** and **one categorical explanatory variable** is sometimes called a "`parallel slopes`" linear regression due to the shape of the predictions.  
- `parallel slopes models` allow for **different intercepts**, but **force all lines to have the same slope**.  

```{r Creating a Parallel Slopes Model}
# Fit a linear regression with a numerical and categorical explanatory variables
mdl_price_vs_both <- lm(price_twd_msq ~ 
                        n_convenience + 
                        house_age_years+0,
                        data = taiwan_real_estate)

# See the result
mdl_price_vs_both
```

```{r Creating a Regreesion Table for Parallel Slopes Model}
get_regression_table(mdl_price_vs_both)
```

## Predicting with Parallel Slopes  
```{r Parallel Slopes Prediction Workflow}
# STEP 1: Make a grid of explanatory data
explanatory_data <- expand_grid(
  # Set n_convenience to zero to ten
  n_convenience = seq(0, 10, 1),
  # Set house_age_years to the unique values of that variable
  house_age_years = unique(taiwan_real_estate$house_age_years)
)

# STEP 2: Add predictions to the data frame
prediction_data <- explanatory_data %>% 
  mutate(
    price_twd_msq = predict(
      mdl_price_vs_both, 
      explanatory_data))

# STEP 3: Create a plot with a layer of predictions.¿
taiwan_real_estate %>% 
  ggplot(aes(n_convenience,
             price_twd_msq,
             color = house_age_years)) +
  geom_point() +
  geom_parallel_slopes(se = FALSE) +
  # Add points using prediction_data, with size 5 and shape 15
  geom_point(
    data = prediction_data,
    size = 5,
    shape = 15
  )
```

## Manually Calculating Predictions  
- For the `parallel slopes` case is that the **intercept is different for each category of the categorical explanatory variable**. That means you need to consider the case when each each category occurs separately.  

```{r Manually Calculating Predictions}
# STEP 1: Get coefficients, slope and intercepts

# Get the coefficients from mdl_price_vs_both
coeffs <- coefficients(mdl_price_vs_both)

# Extract the slope coefficient
slope <- coeffs[1]

# Extract the intercept coefficient for 0 to 15
intercept_0_15 <- coeffs[2]

# Extract the intercept coefficient for 15 to 30
intercept_15_30 <- coeffs[3]

# Extract the intercept coefficient for 30 to 45
intercept_30_45 <- coeffs[4]

# STEP 2: Add columns to explanatory data
# Consider the case when each category occurs separately
prediction_data <- explanatory_data %>% 
  mutate(
    # Consider the 3 cases to choose the intercept
    intercept = case_when(
      house_age_years == "0 to 15" ~ intercept_0_15,
      house_age_years == "15 to 30" ~ intercept_15_30,
      house_age_years == "30 to 45" ~ intercept_30_45
    ),
    
    # Manually calculate the predictions
    price_twd_msq = intercept + slope * n_convenience
  )

# See the results
prediction_data
```

## Assesing Model Performance  
- The `coefficient of determination` or `R-squared`, measures how good the regression's prediction line fits the observed values.  
- The `residual standard error` or `RSE`, is the typical size of the residuals.  
- Including too many explanatory variables can lead to `overfitting`. This means that the model is optimized to **provide the best fit for that particular data set, but no longer reflects the general population**.  
- The `adjusted coefficient of determination` includes a small penalty term for each additional explanatory variable to compensate for this effect.  

```{r Comparing Coefficients of Determination}
mdl_price_vs_conv <- lm(price_twd_msq ~
                        n_convenience,
                        data = taiwan_real_estate)

mdl_price_vs_age <- lm(price_twd_msq ~
                       house_age_years+0,
                       data = taiwan_real_estate)

mdl_price_vs_conv %>% 
  # Get the model-level coefficients
  glance() %>% 
  # Select the coeffs of determination
  select(r.squared, adj.r.squared)

# Get the coeffs of determination for mdl_price_vs_age
mdl_price_vs_age %>% 
  glance() %>% 
  select(r.squared, adj.r.squared)

# Get the coeffs of determination for mdl_price_vs_both
mdl_price_vs_both %>% 
  glance() %>% 
  select(r.squared, adj.r.squared)
```

```{r Comparing Residual Standard Error}

```

# Functions Reference  
- `stats::lm()`: 
- `tidyr::expand_grid()`: 
- `broom::glance()`: retrieves regression model-level metrics as a data frame.  
- `dplyr::pull()`: extracts a specific metric from a regression model.  