---
title: "Intermediate R"
author: "Luis Moreno"
date: "12/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Relational Operators

- The main relational operators are:
  - **equal** with the `==` operator.
  - **not equal** with the `!=` operator.
  - **greater than** another with the `>` operator.
  - **less than** another with the `<` operator.
  - **greater than or equal** to another with the `>=` operator.
  - **less than or equal** to another with the `<=` operator.

- Relational operators can be used with `vectors`, `matrices`, and `lists` objects.

```{r Basic Relational Operators}
# Comparison of logicals
TRUE == FALSE

# Comparison of numerics
-6 * 14 != 17 - 101

# Comparison of character strings
"useR" == "user"

# Compare a logical with a numeric
TRUE == 1
```

## Vectors Comparison
```{r}
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Popular days
linkedin > 15

# Quiet days
linkedin <= 5

# LinkedIn more popular than Facebook
linkedin > facebook
```

## Matrices Comparison
```{r}
# The social data has been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)
views <- matrix(c(linkedin, facebook), nrow = 2, byrow = TRUE)

# When does views equal 13?
views == 13

# When is views less than or equal to 14?
views <= 14
```

# Logical Operators

- The main logical operators are:
  - `AND` or `&`,
  - `OR` or `|`,
  - `NOT` or `!`.

```{r}
# The linkedin and last variable
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
last <- tail(linkedin, 1)

# Is last under 5 or above 10?
last < 5 | last > 10

# Is last between 15 (exclusive) and 20 (inclusive)?
last < 15 & last >= 20
```

```{r}
# Social data
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)
views <- matrix(c(linkedin, facebook), nrow = 2, byrow = TRUE)

# linkedin exceeds 10 but facebook below 10
linkedin > 10 & facebook < 10

# When were one or both visited at least 12 times?
linkedin >= 12 | facebook >= 12

# When is views between 11 (exclusive) and 14 (inclusive)?
views > 11 & views <= 14
```

```{r}
li_df <- read.csv("/Users/luismoreno/Documents/GitHub/datacamp/datasets/li.csv")

# Select the second column, named day2, from li_df: second
second <- li_df$day2

# Build a logical vector, TRUE if value in second is extreme: extremes
extremes <- second > 25 | second < 5

# Count the number of TRUEs in extremes
sum(extremes)
```

# Conditional Statements

```{r if-else statement}
# Variables related to your last day of recordings
most_common_fruit <- "Tomatoes"
my_favorite_fruit <- "Avocadoes"

logical <- most_common_fruit == my_favorite_fruit

# Examine the if statement for medium
if (logical == TRUE) {
  print("Your favorite fruit is the most common fruit")
} else {
  print("Your favorite fruit is NOT the most common fruit")
}
```

```{r if-else if statement}
minimum_age <- 35
your_age <- 36
friends_age <- 42

if (your_age < minimum_age) {
  print("Your age does not satifies the minimum age")
} else if (friends_age < minimum_age) {
  print("Your friend's age does not satifies the minimun age")
} else {
  print("You or your friend satifies the minimum age")
}
```

# Loops

- The `break` statement abandons the active loop: the remaining code in the loop is skipped and the loop is not iterated over anymore.

- The `next` statement skips the remainder of the code in the loop, but continues the iteration.

```{r while loop}
# Initialize the speed variable
speed <- 64

# Code the while loop
while (speed > 30) {
  print("Slow down!")
  speed <- speed - 7
}

# Print out the speed variable
print(speed)
```

```{r loops-break & conditional statements}
# Initialize the speed variable
speed <- 79

while (speed > 30) {
  print(paste("Your speed is", speed))
  
  # Break the while loop when speed exceeds 80
  if (speed > 80) {
    break
  }
  
  if (speed > 48) {
    print("Slow down big time!")
    speed <- speed - 11
  } else {
    print("Slow down!")
    speed <- speed - 6
  }
}
```

```{r}
# Initialize i as 1 
i <- 1

# Code the while loop
while (i <= 10) {
  print(3 * i)
  # Check if i is divisible by 8
  if (8 %% (3 * i) == 0) {
    print(3 * i)
    break
  }
  i <- i + 1
}
```

```{r Loop over a list}
# The nyc list is already specified
nyc <- list(pop = 8405837, 
            boroughs = c("Manhattan", "Bronx", "Brooklyn", "Queens", "Staten Island"), 
            capital = FALSE)

# Loop version 1
for (i in nyc) {
    print(i)
}

# Loop version 2
for (i in 1:length(nyc)) {
    print(nyc[[i]])
}
```

```{r Loop over a matrix}
ttt <- matrix(c("0", NA, "X", NA, "0", "0", "X", NA, "X"), nrow=3, byrow=TRUE)

# define the double for loop
for (i in 1:nrow(ttt)) {
  for (j in 1:ncol(ttt)) {
    print(paste("On row",i,"and column",j,"the board contains",ttt[i,j]))
  }
}
```

# Functions
- To consult the `documentation` on a function use the following R commands:
  - `help(function_name)`
  - `?function_name`

- To consult the `arguments` of a function use the following R command:
  - `args(function_name)`

- The `...` on function arguments is called the `ellipsis`, and it is a way to pass arguments along without the function having to name them explicitly.

- Function arguments that have a default value makes them **optional**.

